---
title: "Proj1"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(magrittr)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)

```
## Part 1: Data scraping and preparation
## Step 1: Scrape your competitor's data
This pipeline first reads the html from the provided url and finds the table based on the HTML node. Column names are added by using a vectpr of column names and the pipeline results in a dataframe
```{r step1, include=TRUE}

url <- "https://www.spaceweatherlive.com/en/solar-activity/top-50-solar-flares"

sf <- url %>%
  read_html() %>%
  html_node(".table-striped") %>%
  html_table() %>%
  set_colnames(c("rank", "flare_classification","date","flare_region", "start_time","maximum_time","end_time","movie")) %>%
  as.data.frame()
head(sf, n = 10)

```
## Step 2: Tidy the top 50 solar flare data
uses the data frame from Step 1 and first drops the movie column by using the select function that keeps everything but the movie col using the minus sign operator. Then the unite function is used to create a new column based on the column vectors passed in and finally the united column is converted to a POSIXct type using type convert and using a column specification which uses a regex.
```{r step2, include=TRUE}
sf <- sf %>%
  select(-movie) %>%
  unite ("start_datetime", c("date", "start_time"), sep = " ", remove =    FALSE)%>%
  unite ("max_datetime",  c("date", "maximum_time"),sep = " ", remove = FALSE)%>%
  unite ("end_datetime", c("date", "end_time"), sep = " ", remove = TRUE)%>%
 type_convert(cols(start_datetime=col_datetime(format = "%Y/%m/%d %H:%M"),   max_datetime=col_datetime(format = "%Y/%m/%d %H:%M"), end_datetime = col_datetime(format = "%Y/%m/%d %H:%M")))
head(sf, n = 10)
```


## Step 3: Scrape the NASA data
```{r step3, include=TRUE}
url <-  "http://cdaw.gsfc.nasa.gov/CME_list/radio/waves_type2.html"

NASA <- url %>%
  read_html() %>%
  html_node("pre")%>%
  html_text() %>%
  strsplit("\n") %>%
  as.data.frame() %>%
  set_colnames(c("col")) %>%
  separate(col, into = c("start_date", "start_time", "end_date", "end_time", "start_freq", "end_freq", "Location", "NOAA", "Importance", 
                      "CME_date", "CME_time", "CME_CPA", "CME_width", "CME_speed", "Plots"), sep = "[ ]+")

tail(NASA, 513) %>%
  head(511)
```
## Step 4: Tidy the NASA Table
```{r step4, include = TRUE}
NASA[NASA == "------"] <- NA
NASA[NASA == "-----"] <- NA
NASA[NASA == "----"] <- NA
NASA[NASA == "--/--"] <- NA
NASA[NASA == "--:--"] <- NA
NASA$Halo <- NA
NASA[c("start_date", "start_time", "end_date", "end_time", "start_freq", "end_freq", "Location", "NOAA", "Importance", 
                      "CME_date", "CME_time", "CME_CPA", "CME_width", "CME_speed", "Plots", "Halo")]

NASA$Halo[NASA$CME_CPA == "Halo"] <- "Yes"
NASA$CME_CPA[NASA$CME_CPA == "Halo"] <- NA
NASA[c("Halo")][is.na(NASA[c("Halo")])] <- "No"
NASA$Lower_bound <- NA
NASA[c("start_date", "start_time", "end_date", "end_time", "start_freq", "end_freq", "Location", "NOAA", "Importance", 
                      "CME_date", "CME_time", "CME_CPA", "CME_width", "CME_speed", "Plots", "Halo", "Lower_bound")]
NASA$Lower_bound[grepl(">", NASA$CME_width)] <- "Yes"
NASA[c("Lower_bound")][is.na(NASA[c("Lower_bound")])] <- "No"
NASA$CME_width[grepl(">", NASA$CME_width)] <- NASA$CME_width
NASA <- tail(NASA, 513) %>%
  head(511)
NASA
```


